---
title: "QMB Exercise 1 - Exploring Housing Rents"
author: "Dino Nienhold"
date: "Wednesday, February 18, 2015"
output: pdf_document
---

#Introduction
The following report is based on the QMB Exercise 1 - Exploring Housing Rents. The task description pdf file is bis_ex1-HousingRents.pdf

#Requirements
Please make sure that you the following packages loaded in your workspace.
```{r}
library("dplyr")
library("ggplot2")
```

#Data Set
Please make sure you have the file housingrents.csv in the subdirectoy Data in your workspace.
```{r}
housingrents <- read.csv("./Data/housingrents.csv",sep=";")
```

#Task 1
There are `dim(housingrents)[1]` observations and `dim(housingrents)[2]` variables in the dataset.
The str command gives an overview of the variable types:
```{r}
str(housingrents)
```
There are 14 NA values in the balcony variable.
```{r}
summary(housingrents)
```
#Task 2

##Data Processing
For analysis purposes it is necessary to convert the rooms and nre variable to a factor.
```{r}
housingrents <- mutate(housingrents, rooms = factor(rooms), nre = factor(nre,levels=c(0,1),labels=c("no","yes")))
```

##Plotting
The following chart shows the frequency of appartments according to their numberof rooms.
```{r}
ggplot(data=housingrents, aes(x=rooms,label=rooms))  + geom_bar(fill="blue") + ggtitle("Count by Rooms") 
```

#Task 3
```{r}
rooms2nre <- xtabs(~rooms+nre, data=housingrents)
rooms2nre <- prop.table(rooms2nre,1)
ggplot(data.frame(rooms2nre), aes(x=rooms, y=Freq, fill=nre)) + geom_bar(stat="identity") + 
  xlab("Number of Rooms") + ylab("Ratio of NRE Property to non-Property") + 
  ggtitle("NRE Property by Number of Rooms") 
addmargins(prop.table(rooms2nre,1))
```

http://stats.stackexchange.com/questions/798/calculating-optimal-number-of-bins-in-a-histogram-for-n-where-n-ranges-from-30
```{r}
bw <- diff(range(housingrents$area)) / (2 * IQR(housingrents$area) / length(housingrents$area)^(1/3))

cfunc <- function(x, n) (sqrt(n) * (mean(x) - (1/lambda))) / (1/lambda)
dat <- data.frame(x = apply(data, 2, cfunc, 40),size = factor(rep(c(40), rep(numberSimulation, 1))))
g <- ggplot(housingrents, aes(x = area)) + geom_histogram(binwidth = bw,colour="black", fill="white")  + geom_density(alpha=.2) + geom_density(alpha=.5, fill="#FF6666") +
    geom_vline(aes(xintercept=mean(area, na.rm=T)), color="red", linetype="dashed", size=1)
g <- ggplot(data.frame(x), aes(x = data.frame(x)[1])) + geom_histogram() 
g <- g + stat_function(fun = dnorm, size=50)
g + facet_grid(. ~ size)
qqnorm(housingrents$area);qqline(housingrents$area, col = 2)
boxplot(housingrents$area)
```